Viscous Transport in Eroding Porous Media

### 2019-09-09 docs/transport/erosion_transport.tex
working through some of the results

In Figure~\ref{fig:ErodingLow20Transport}(b), we plot the temporal
evolution of the particle spreading $\sigma_\lambda$.  As in the last
example, we analyze the spreading at several different porosities and we
use the reinsertion algorithm described in Section~\ref{sec:dispersion}
to form sufficiently long trajectories to observe an asymptotic
dispersion rate.  For all the porosities, the dispersion is much closer
to ballistic when compared to the results in
Figure~\ref{fig:Eroding20Transport}.  However, there are still clear
transitions from ballistic dynamics at early times to asymptotic
super-diffusive dynamics.  In contrast to the lower porosity initial
condition (Section~\ref{sec:Eroding20}), at early times the erosion
results in a decrease in particle spreading. In particular, once erosion
has eroded the first 5\% of the bodies, the particle spreading
transitions from $\sigma_\lambda \sim t^{0.92}$ to $\sigma_\lambda \sim
t^{0.69}$.  To explain this behavior, note that the velocity is fairly
uniform, albeit small, at the initial configuration (see
Figure~\ref{fig:ErodingLow20vort}).  Then as the geometry erodes, the
ratio of the fast and slow regions increases.  Since super-diffusive
dispersion results from a tracers spending times in a both slow and fast
regions, this results in an increase in the rate of super-dispersion.
\todo[inline]{Still need to work on this and explain how it contrasts
from the other 20 body example}
\todo[inline]{Need a reference for last sentence}

@@

\todo[inline]{do better job of explaining why diesperions does what it
does.  For this problem, there are lots more of small bodies rather than
a few big ones.  Therefore, after reinsertion, many of the bodies are
placed in a ballistic regime which means that anoamlous dispersion is no
longer observed.}
Next, we use the 1000 tracer trajectories to compute the second
(Figure~\ref{fig:Eroding100Transport}(b)) moments of the streamlines at
various porosities.  At high porosities, the first moment is similar to
the first-moment of trajectories in a empty channel (black dashed line).
Similar to the other previous examples, we initially observe ballistic
motion, and then super-diffusive dispersion.  However, similar to the
example in Figure~\ref{fig:ErodingLow20Transport}(b), the anomalous
dispersion rate is not growing with the porosity.  Therefore, it appears
that the dispersion rate in an eroding geometry depends not only the
porosity, but also the location and shape of the bodies.

@@

Finally, using network models for flow in porous media~\cite{}, the
anomalous dispersion rate can be linked to channel widths between solid
bodies.  As described in Section~\ref{sec:throats}, we use a Delaunay
triangulation to define neighboring eroding bodies, and we compute the
throat size by finding the closest distance between all neighboring
bodies.  Once all throat sizes are computed, we analyze their
distribution as a function of the porosity.  In
Figure~\ref{fig:Eroding100gap_hist}, we plot histograms of the throat
sizes at six porosities throughout the erosion process.

\begin{figure}[H]
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_1}
\caption{}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_46}
\caption{}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_112}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_164}
\caption{}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_207}
\caption{}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_246}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap_hist} The distribution of the gaps in the
case of erosion of 100 bodies at six porosities and the fitting curves of Weibull distribution 
described by the shape parameter $k > 0$ and the scale parameter $\lambda > 0$. 
(a) 100 bodies with 258 gaps as porosity $\phi = 50.09 \%$ 
and $(k, \lambda)=(1 .4650, 0.0605)$ for Weibull distribution. 
(b) 100 bodies with 258 gaps as porosity $\phi = 55.02 \%$ 
and $(k, \lambda)=( 1.9485, 0.0737)$ for Weibull distribution. 
(c) 97 bodies with 249 gaps as porosity $\phi = 65.03 \%$ 
and $(k, \lambda)=( 2.5682, 0.0965)$ for Weibull distribution. 
(d) 92 bodies with 234 gaps as porosity $\phi = 75.14 \%$ 
and $(k, \lambda)=( 2.7771, 0.1255)$ for Weibull distribution. 
(e) 75 bodies with 189 gaps as porosity $\phi = 85.07 \%$ 
and $(k, \lambda)=( 2.6235, 0.1713)$ for Weibull distribution. 
(f) 34 bodies with 74 gaps as porosity $\phi = 95.12 \%$ 
and $(k, \lambda)=( 1.9840, 0.3194)$ for Weibull distribution. }
\end{figure}

In Figure~\ref{fig:Eroding100gap_mean_var},
we plot the mean and variance of the throat sizes as a function of the
porosity.  While the mean steadily grows with the porosity, the variance
remains relatively flat for a wide range of porosities.

In Figure~\ref{fig:Eroding100gap_mean_var}, 
we record the gap distances between bodies neighbors and
calculate the mean and variance with respect to porosities. 
For the mean of gap distances, it is linearly increacing until $\phi = 85\%$ 
which means the eroding . 
For the variance of gap distances, it is almost a constant function until $\phi= 85\%$  
\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_mean}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height=0.8\linewidth]{./figs/gap_variance}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap_mean_var}  The mean of gaps in (a) and
the variance with of gaps in (b) with respect to porosity $\phi$.  Red
stars are six selected porosities to have the details of  the
distribution of gaps and fitting curves of the distributions in
Figure~\ref{fig:Eroding100gap_hist}.}
\end{figure}

We plot the distribution of the gap sizes at six porosities
in Figure~\ref{fig:Eroding100gap_hist}.  At the initial configuration, there
are 100 bodies and 258 pairs of neighbors.  The mean distance between
these neighbors is $???  \times 10^{-???}$ and the variance is $???
\times 10^{???}$.  As the bodies erode, the mean and variance of the
distance between the bodies increases.  Finally, when the porosity reaches $95.12\%$
there are only 34 bodies and 74 pairs of neighbors.

%\begin{figure}[H]
%\begin{center}
%\includegraphics[width = 0.8\textwidth]{./figs/100b_t100tracer}
%\caption{\label{fig:Eroding100tracer} The historical trajectories of 
%200 tracers released at x = -1 in the fluid with multiple bodies.}
%\end{center}
%\end{figure}

@@

%\begin{figure}[H]
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/tort_eulerian20b}
%\caption{\label{fig:ErodingLow20tort} The tortuosity inside of an
%eroding geometry initialized with 20 bodies.  The tortuosity is
%calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue
%dots).  The dash line is the fitting line $\widehat{T}(\phi)=\phi^{-p}$
%as $p=0.1669$ and the root mean square deviation is $1.13 \times
%10^{-2}$.}
%\end{figure}
%
%\begin{figure}[H]
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/20b_dense_second_moment_ref}
%\caption{\label{fig:Eroding20anomalousLow} The analysis of anomalous
%dispersion in 20 nearly touching bodies with respect to initial porosity
%(as p=30.67\%) and another seven porosities during the eroding. The
%dashed line correspondes to a porosity of 100\% and the dashed-dotted
%lines are lines of best fit with slope 0.9194 (as p=95\%), 0.8666 (as
%p=85.17\%), 0.8682 (as p=75.10\%), 0.9087 (as p=65.02\%), 0.9054 (as
%p=55.08\%), 0.7234 (as p=45.03\%), 0.6895 (as p=35.05\%), and 0.9225 (as
%p=30.67\%).}
%\end{figure}
%
%\begin{figure}[H]
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/tort_eulerian100}
%\caption{\label{fig:Eroding100tort_all} The Eulerian and Lagrangian
%approach of tortuosity.  The blue dots are Eulerian approach of
%tortuosity and the red stars and square are Lagrangian approach of
%tortuosity.  The dash line is the fitting line
%$\widehat{T}(\phi)=\phi^{-p}$ as $p=0.2459$ and the root-mean-square
%error is $5.5 \times 10^{-3}$.  We note that the model
%$\widehat{T}(\phi) = 1-p\ln(\phi)$, with $p=0.2631$ has a comparable
%root-mean-square error of $5.2 \times 10^{-3}$.}
%\end{figure}
%
%\begin{figure}
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/100b_second_moment_long_ref}
%\caption{\label{fig:Eroding100anomalous} The analysis of anomalous
%dispersion in 100 bodies with respect to initial porosity
%(as p=50.09\%) and another five porosities during the eroding. The dashed line
%correspondes to a porosity of 100\% and the dashed-dotted lines are lines
%of best fit with slope 1.1135 (as p=95.12\%), 0.6825 (as p=85.07\%), 
%0.7347 (as p=75.14\%), 0.5915 (as p=65.03\%), 
%0.6107 (as p=55.02\%), and 0.7 (as p=50.09\%).}
%\end{figure}

>>>>

In Figure~\ref{fig:ErodingLow20Transport}(b), we plot the temporal
evolution of the particle spreading $\sigma_\lambda$.  As in the last
example, we analyze the spreading at several different porosities and we
use the reinsertion algorithm described in Section~\ref{sec:dispersion}
to form sufficiently long trajectories to observe an asymptotic
dispersion rate.  For all the porosities, the dispersion is much closer
to ballistic when compared to the results in
Figure~\ref{fig:Eroding20Transport}.  However, there are still clear
transitions from ballistic dynamics to asymptotic super-diffusive
dynamics.  In contrast to the lower porosity initial condition
(Section~\ref{sec:Eroding20}), at early times the erosion results in a
decrease in dispersion rate. In particular, once erosion has eroded the
first 5\% of the bodies, the particle spreading transitions from
$\sigma_\lambda \sim t^{0.92}$ to $\sigma_\lambda \sim t^{0.69}$.  To
explain this behavior, recall that anomalous dispersion is caused by
tracers spending time in both the fast and slow regimes.  Since the
initial configuration has a reasonably uniform velocity (see
Figure~\ref{fig:ErodingLow20vort}), albeit a small one, the dispersion
is nearly ballistic. However, as the geometry erodes, the flow becomes
more intermittent, and this results in an increased anomalous dispersion
rate~\cite{dea-leb-den-tar-bol-dav2013}.  Then, as the bodies continue
to erode, the geometry channelizes, and most tracers are transported
with a large velocity through the channels, again resulting in a nearly
ballistic motion~\cite{sie-ili-pri-riv-gua2019}.

@@

In Figure~\ref{fig:Eroding100Transport}(b), we use the 1000 tracer
trajectories to compute the asymptotic dispersion rates at various
porosities.  Again, we initially observe ballistic motion (black dashed
line), and then super-diffusive dispersion.  Similar to the example in
Figure~\ref{fig:ErodingLow20Transport}(b), the anomalous dispersion rate
is not growing with the porosity.  Therefore, it appears that the
dispersion rate in an eroding geometry depends not only the porosity,
but also the location and shape of the bodies.  Finally, at the highest
porosity, anomalous dispersion is only observed briefly in the time
interval $(0.5,1)$, and then transitions back to a ballistic regime.
Since the bodies are so small at this high porosity, after reinsertion,
the streamline is not significantly deflected by any of the bodies, and
this results in a ballistic regime.

@@

Finally, we investigate the effects of erosion on pore sizes.  The
distribution of the pore sizes can be directly related to the
distribution of the velocity, and thus effects the
tortuosity~\cite{den-ica-hid2018} and anomalous
dispersion~\cite{dea-qua-bir-jua2018}. In addition, the pore sizes are
required in network models~\cite{bry-mel-cad1993, bry-kin-mel1993}.  As
described in Section~\ref{sec:throats}, we use a Delaunay triangulation
to define neighboring eroding bodies, and we compute the throat size by
finding the closest distance between all neighboring bodies.  Once all
throat sizes are computed, we analyze their distribution as a function
of the porosity.

In Figure~\ref{fig:Eroding100gap_hist}, we plot histograms of the pore
throat sizes at six porosities of the erosion process. These
distributions have been modelled with Weibull
distributions~\cite{ioa-cha1993}, and we superimpose the Weibull
distribution with the same first two moments as the data.  The
parameters that parameterize the distribution, $(k,\lambda)$, are
included in the caption of Figure~\ref{fig:Eroding100gap_hist}.  In
Figure~\ref{fig:Eroding100gap_mean_var}, we plot the mean and variance
of the throat sizes as a function of the porosity.  At low porosities,
the mean grows roughly linearly and the variance is roughly constant
until about porosity $\phi = 0.85$.  At this point, both the mean and
variance of the gap sizes quickly grow.

\begin{figure}[H]
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_1}
\caption{100 bodies, 258 gaps}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_46}
\caption{100 gaps, 258 gaps}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_112}
\caption{97 bodies, 249 gaps}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_164}
\caption{92 bodies, 234 gaps}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_207}
\caption{75 bodies, 189 gaps}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_246}
\caption{34 bodies, 74 gaps}
\end{subfigure}
\caption{\label{fig:Eroding100gap_hist} The distribution of the gaps in
the case of 100 bodies eroding at six porosities. The black curves are
Weibull distributions whose first and second moments agree with the
data. The porosities and Weibull distribution parameters $(k,\lambda)$
at each of the porosities are: (a) $\phi = 50.09\%$ and
$(k,\lambda)=(1.4650,0.0605)$; (b) $\phi = 55.02\%$ and
$(k,\lambda)=(1.9485,0.0737)$; (c) $\phi = 65.03\%$ and
$(k,\lambda)=(2.5682,0.0965)$; (d) $\phi = 75.14\%$ and
$(k,\lambda)=(2.7771,0.1255)$; (e) $\phi = 85.07\%$ and
$(k,\lambda)=(2.6235,0.1713)$; (f) $\phi = 95.12\%$ and
$(k,\lambda)=(1.9840, 0.3194)$.}
\end{figure}

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_mean}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height=0.8\linewidth]{./figs/gap_variance}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap_mean_var} The mean (a) and variance
(b) of the gap sizes of an eroding geometry.  The red stars correspond
to the six porosities in Figure~\ref{fig:Eroding100gap_hist}.}
\end{figure}

@@

%\begin{figure}[H]
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/tort_eulerian20b}
%\caption{\label{fig:ErodingLow20tort} The tortuosity inside of an
%eroding geometry initialized with 20 bodies.  The tortuosity is
%calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue
%dots).  The dash line is the fitting line $\widehat{T}(\phi)=\phi^{-p}$
%as $p=0.1669$ and the root mean square deviation is $1.13 \times
%10^{-2}$.}
%\end{figure}
%
%\begin{figure}[H]
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/20b_dense_second_moment_ref}
%\caption{\label{fig:Eroding20anomalousLow} The analysis of anomalous
%dispersion in 20 nearly touching bodies with respect to initial porosity
%(as p=30.67\%) and another seven porosities during the eroding. The
%dashed line correspondes to a porosity of 100\% and the dashed-dotted
%lines are lines of best fit with slope 0.9194 (as p=95\%), 0.8666 (as
%p=85.17\%), 0.8682 (as p=75.10\%), 0.9087 (as p=65.02\%), 0.9054 (as
%p=55.08\%), 0.7234 (as p=45.03\%), 0.6895 (as p=35.05\%), and 0.9225 (as
%p=30.67\%).}
%\end{figure}
%
%\begin{figure}[H]
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/tort_eulerian100}
%\caption{\label{fig:Eroding100tort_all} The Eulerian and Lagrangian
%approach of tortuosity.  The blue dots are Eulerian approach of
%tortuosity and the red stars and square are Lagrangian approach of
%tortuosity.  The dash line is the fitting line
%$\widehat{T}(\phi)=\phi^{-p}$ as $p=0.2459$ and the root-mean-square
%error is $5.5 \times 10^{-3}$.  We note that the model
%$\widehat{T}(\phi) = 1-p\ln(\phi)$, with $p=0.2631$ has a comparable
%root-mean-square error of $5.2 \times 10^{-3}$.}
%\end{figure}
%
%\begin{figure}
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/100b_second_moment_long_ref}
%\caption{\label{fig:Eroding100anomalous} The analysis of anomalous
%dispersion in 100 bodies with respect to initial porosity
%(as p=50.09\%) and another five porosities during the eroding. The dashed line
%correspondes to a porosity of 100\% and the dashed-dotted lines are lines
%of best fit with slope 1.1135 (as p=95.12\%), 0.6825 (as p=85.07\%), 
%0.7347 (as p=75.14\%), 0.5915 (as p=65.03\%), 
%0.6107 (as p=55.02\%), and 0.7 (as p=50.09\%).}
%\end{figure}
%
%\begin{figure}[H]
%\begin{center}
%\includegraphics[width = 0.8\textwidth]{./figs/100b_t100tracer}
%\caption{\label{fig:Eroding100tracer} The historical trajectories of 
%200 tracers released at x = -1 in the fluid with multiple bodies.}
%\end{center}
%\end{figure}
